---
output:
  pdf_document: default
  html_document: default
---

# Problem Sheet 1 - Solutions {-}

### Question 0 * {-}

Give an example of each of the following data types:

i. Categorical

__SOLUTION:__

Dog or Cat

ii. Interval

__SOLUTION:__

Height

iii. Ordinal

__SOLUTION:__

What did you think of the first class:

Hated it,  Neutral, Loved it



## Measures of Location {-}
### Question 1 {-}

Describe each of the following measures of location including their the pros and cons

i. Mean

The mean is the point at which the sum of the deviations is 0:

Pros:


* Easy to calculate;

* Uses all the data.

Con:

* Sensitive to extreme values.
 


ii. Median

The median is the middle value of the ordered set of values:

Pros:

* Median gives the center of the data;

* Not sensitive to extreme values.

Con:

* Does not use all the data.
	 


iii. Variance

The variance is the spread around the mean:

Pro:

* Takes all data into account.

Cons:


* Hard to interpret;

* Can be influenced by extreme values.
	


iv. Skewness


Skewness is a measure of symmetry of a distribution:
Pro

* Takes all data into account;

Con

* Not intuitive.


## Counting Questions  {-}
### Question 2 {-}

In a given card game hand each player gets four cards

i.	How many different combinations of four cards can be made for a 52 card deck, when __order does__ matter.



__SOLUTION:__


$$P_4^{52}=\frac{52!}{(52-4)!}=(52\times51\times50\times49)=6,497,400 $$

```{r P1aQ2a,echo=FALSE}
Q2a<-52*51*50*49

# OR
Q2a<-factorial(52)/factorial(48)

```


ii.	How many different combinations of four cards can be made for a 52 card deck, when order __does not__ matter.


__SOLUTION:__
```{r P1aQ2b,echo=FALSE}
Q2b<-52*51*50*49/(4*3*2*1)

# OR
Q2b<-choose(52,4)

```

Order __does not__ matter
$$C_4^{52}=\big(\begin{array}{c} 52\\ 4 \end{array} \big)=\frac{52!}{(52-4)!4!}=\frac{(52\times51\times50\times49)}{(4\times3\times2\times1)}=`r Q2b` $$



iii.	What is the probability of being dealt a hand with the 2 of clubs, the 3 of clubs, the 4 of clubs and the 5 of clubs, __in exact order__?

__SOLUTION:__

$$\frac{1}{P_4^{52}}=\frac{1}{\frac{52!}{(52-4)!}}=\frac{1}{(52\times51\times50\times49)}=`r 1/Q2a` $$



iv.	What is the probability of being dealt a hand with the 2 of clubs, the 3 of clubs, the 4 of clubs and the 5 of clubs, __in any order__?

__SOLUTION:__

$$\frac{1}{C_4^{52}}=\frac{1}{\frac{52!}{(52-4)!4!}}=\frac{4\times3\times2\times1}{(52\times51\times50\times49)}=`r 1/Q2b` $$

v.	What is the probability of being dealt a hand with the 2 of diamonds, the 7 of spades, the 9 of clubs and the Queen of diamonds, in any order?

__SOLUTION:__

$$\frac{1}{C_4^{52}}=\frac{1}{\frac{52!}{(52-4)!4!}}=\frac{4\times3\times2\times1}{(52\times51\times50\times49)}=`r 1/Q2b` $$


vi.	What is the probability of being dealt a hand with at least one King, in any order?

__SOLUTION:__

```{r P1aQ2f,echo=FALSE}
Q2fi<-4*48*47*46/(1*3*2*1)
Q2fii<-4*3*48*47/(2*1*2*1)
Q2fiii<-4*3*2*48/(3*2*1*1)
Q2fiv<-1
Q2fo<-48*47*46*45/(4*3*2*1)
Q2fi+Q2fii+Q2fiii+Q2fiv+Q2fo
```

Number of hands with at least one king in any order:
Hands with exactly One King

$$\frac{(4\times48\times47\times46)}{(1\times3\times2\times1)}=`r Q2fi` $$
Hands with exactly Two Kings

$$\frac{(4\times3\times48\times47)}{(2\times1\times2\times1)}=`r Q2fii` $$
Hands with exactly Three Kings

$$\frac{(4\times 3 \times 2\times48)}{(3\times 2\times1\times1)}=`r Q2fiii` $$

Hands with exactly Four Kings

$$\frac{(4\times 3 \times 2\times1)}{(4\times 3\times 2\times1\times1)}=`r Q2fiv` $$
Hands with exactly Zero Kings

$$\frac{(48\times 47 \times 46\times45)}{(4\times3\times 2\times1)}=`r Q2fo` $$
All possible hands
`r Q2fo+Q2fi+Q2fii+Q2fiii+Q2fiv`
`r Q2b`

At least one King is all possible hands minus no kings
`r Q2fi+Q2fii+Q2fiii+Q2fiv`
$$\frac{\frac{(4\times51\times50\times49)}{(1\times3\times2\times1)}}{\frac{52!}{(52-4)!4!}}=\frac{4\times51\times50\times49}{(52\times51\times50\times49)}=`r (Q2fi+Q2fii+Q2fiii+Q2fiv)/Q2b` $$


### Question 3  {-}

A bank issues bank cards with PINs consisting of 4 digits, each one {0,1,2,…,9}. 

A. Given that any 4-digit code can be used.
i.	How many unique PINs can be generated?


__SOLUTION:__

```{r P1aQ3i,echo=FALSE}
Q3ia<-10*10*10*10
```


$$10^4=10\times 10 \times 10\times 10\times 10=`r Q3ia`$$

ii.	What is the probability of guessing someone’s PIN?

__SOLUTION:__

$$\frac{1}{10^4}=\frac{1}{10\times 10 \times 10\times 10\times 10}=`r 1/Q3ia`$$

iii.	What is the probability of guessing someone’s PIN given you know the first number?

__SOLUTION:__

$$\frac{10}{10^4}=\frac{10}{10\times 10 \times 10\times 10\times 10}=`r 10/Q3ia`$$

B. The digits must be different.

i.	How many unique PINs can be generated?

__SOLUTION:__


Order __does__ matter
$$P_4^{10}=\frac{10!}{(10-4)!}=(10\times9\times8\times7)=5,040 $$
```{r P1aQ3ii,echo=FALSE}
Q3iia<-10*9*8*7
```

ii.	What is the probability of guessing someone’s PIN?

__SOLUTION:__

$$\frac{1}{P_4^{10}}=\frac{1}{\frac{10!}{(10-4)!}}=\frac{1}{(10\times9\times8\times7)}=`r 1/Q3iia` $$

iii.	What is the probability of guessing someone’s PIN give you know the first number?

__SOLUTION:__

$$\frac{10}{P_4^{10}}=\frac{10}{\frac{10!}{(10-4)!}}=\frac{10}{(10\times9\times8\times7)}=`r 10/Q3iia` $$


### Question 4   {-}

 In a lottery, each ticket has 5 one-digit numbers 0-9 which is not repeated on it.
 
A. You win if your ticket has the digits in any order. 

i.	What are the total number of possible combinations?



__SOLUTION:__


Order __does not__ matter
$$C_5^{10}=\big(\begin{array}{c} 10\\ 5 \end{array} \big)=\frac{10!}{(10-5)!5!}=\frac{(10\times9\times8\times7\times6)}{(5\times4\times3\times2\times1)}=252 $$
```{r P1aQ4i,echo=FALSE}
Q4i<-choose(10,5)
Q4i<-10*9*8*7*6/(5*4*3*2*1)

```


ii.	What is the probability of winning?

__SOLUTION:__


Order __does not__ matter
$$\frac{1}{\big(\begin{array}{c} 10\\ 5 \end{array} \big)}=\frac{1}{\frac{10!}{(10-5)!5!}}=\frac{(5\times4\times3\times2\times1)}{(10\times9\times8\times7\times6)}=`r 1/Q4i` $$

B. You would win only if your ticket has the digits in the required order.

i.	What are the total number of possible combinations?


__SOLUTION:__


Order __does__ matter
$$P_5^{10}=\frac{10!}{5!}=10\times9\times8\times7\times6=30240 $$
```{r P1aQ4ii,echo=FALSE}
Q4ii<-10*9*8*7*6

```

ii.	What is the probability of winning?

$$\frac{1}{P_5^{10}}=\frac{1}{\frac{10!}{5!}}=\frac{1}{10\times9\times8\times7\times6}=`r 1/Q4ii` $$


### Question 5  {-}

How many different combinations of 6 cards can be made for a 52 card deck if

i. order matters.


__SOLUTION:__


Order __does__ matter
$$P_6^{52}=\frac{52!}{(52-6)!}=52\times51\times50\times49\times48\times47=14,658,134,400 $$
```{r P1aQ5i,echo=FALSE}
A<-52*51*50*49*48*47
A<-factorial(52)/factorial(46)
```

ii. order does not matter

__SOLUTION:__


Order __does not__ matter
$$C_6^{52}=\big(\begin{array}{c} 52\\ 6 \end{array} \big)=\frac{52!}{(52-6)!6!}=\frac{(52\times51\times50\times49\times48\times47)}{(6\times5\times4\times3\times2\times1)}=20,358,520 $$
```{r P1aQ5ii,echo=FALSE}
A<-choose(52,6)
A<-52*51*50*49*48*47/(6*5*4*3*2*1)

```

### Question 6  {-}

In a game of 5 card poker what are the number of different possible hands are there?

i.	What is the probability of being dealt a hand with a pair?


__SOLUTION:__

Number of hands with a pair (or more):
$$\frac{52\times3}{2\times1}\times\frac{(48\times47\times46)}{(3\times2\times1)}=1,349,088$$
```{r P1aQ7i,echo=FALSE}
Q7i<-(52*3)/(2*1)*48*47*46/(3*2*1)
Q7_all<-choose(52,5)

```

Number of hands all possible hands
$$\frac{(52\times 51\times 50 \times49\times 48)}{(5\times4\times3\times2\times1)}=`r Q7_all`$$
Number of all possible hands:

$$p(pair)=\frac{\frac{52\times3}{2\times1}\times\frac{(48\times47\times46)}{(3\times2\times1)}}{\frac{(52\times 51\times 50 \times49\times 48)}{(5\times4\times3\times2\times1)}}=`r Q7i/Q7_all`$$

(Alternative) A hand with a only pair

__SOLUTION:__

```{r P1aQ7a,echo=FALSE}
Q7aa<-(52*3)/(2*1)*48*44*40/(3*2*1)

```

$$\frac{52\times3}{2\times1}\times\frac{(48\times44\times40)}{(3\times2\times1)}=`r Q7aa`$$


$$p(pair)=\frac{\frac{52\times3}{2\times1}\times\frac{(48\times44\times40)}{(3\times2\times1)}}{\frac{(52\times 51\times 50 \times49\times 48)}{(5\times4\times3\times2\times1)}}=`r Q7aa/Q7_all`$$


ii.	What is the probability of being dealt a hand with two pair?


__SOLUTION:__

```{r P1aQ7b,echo=FALSE}
Q7b<-((52*3)/(2*1)*48*3/(2*1))/2*44/1
```

Number of two pair hands:

$$\frac{\frac{52\times3}{2\times1}\times\frac{48\times3}{2\times1}}{2\times1}\times\frac{44}{1}=`r Q7b`$$



$$p(\text{two pair})=\frac{\frac{\frac{52\times3}{2\times1}\times\frac{48\times3}{2\times1}}{2\times1}\times\frac{44}{1}}{\frac{(52\times 51\times 50 \times49\times 48)}{(5\times4\times3\times2\times1)}}=`r Q7b/Q7_all`$$


iii.	What is the probability of being dealt a hand with three of a kind and two of kind (Full House)

__SOLUTION:__

Numer of Full House hands:

$$\frac{52\times3\times2}{3\times2\times1}\times\frac{48\times3}{2\times1}=3,744$$
```{r P1aQ7c,echo=FALSE}
Q7c<-((52*3*2)/(3*2*1)*48*3/(2*1))
```

$$p(\text{Full House})=\frac{\frac{52\times3\times2}{3\times2\times1}\times\frac{48\times3}{2\times1}}{\frac{(52\times 51\times 50 \times49\times 48)}{(5\times4\times3\times2\times1)}}=`r Q7c/Q7_all`$$


iv.	What is the probability of being dealt a hand with a Flush (all the same suit)

__SOLUTION:__

```{r P1aQ7d,echo=FALSE}
Q7d<-((52*12*11*10*9)/(5*4*3*2*1))
```

Number of possible flush hands 
$$\frac{52\times12\times11\times10\times9}{5\times4\times3\times2\times1}=`r Q7d`$$

$$p(Flush)=\frac{\frac{52\times12\times11\times10\times9}{5\times4\times3\times2\times1}}{\frac{(52\times 51\times 50 \times49\times 48)}{(5\times4\times3\times2\times1)}}=`r Q7d/Q7_all`$$


## Multiple-Choice Questions {-}

### MCQ Question 7 {-}
Which of the following is an example of an ordinal data type?

A. Temperature in Celsius

B. Eye color

C. Customer satisfaction rating (e.g., Poor, Fair, Good, Excellent)

D. Bank account number

Answer: C. Customer satisfaction rating

### MCQ Question 8 {-}
Which measure of location is least affected by extreme values?

A. Mean

B. Median

C. Mode

D. Variance

Answer: B. Median

### MCQ Question 9 {-}
How many different ways can the letters of the word "STATISTICS" be arranged?

A) 3,628,800

B) 50,400

C) 63,504

D) 1,260

__ANSWER: B) 50,400__

Explanation: 
This is a permutation problem with repeated elements. The total number of arrangements is given by 10!/(3!3!2!)

### MCQ Question 10 {-}

You have 10 different players and need to form a lineup of 4 players, where the order in which they are arranged does not matter. What rule should you use to calculate how many different lineups can be made?

A) Combination

B) Permutation

C) Neither


__ANSWER: A) __

Explanation:
This is a combination problem where order does not matter. 

### MCQ Question 11 {-}

What rule would you use to calculate  how many ways can you arrange the letters in the word "BALLOON"


A) Combination

B) Permutation

C) Neither


__ANSWER: C) __

Explanation:
As there are two L's and O's you have to take that into account


### Question 12 {-}

Write your own question or multiple choice question with solution on counting.

