# Tutorial 4 Solutions {-}

## Named Discrete Distributions {-}

```{r , include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE, fig.align = "center", fig.pos = "H",dpi = 300)
```


```{r include=FALSE,echo=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
library(ggpubr)
library(knitr)
library(ggplot2)
library(ggpubr)
library(knitr)
library(rstatix)
library(tidyverse)
library(dplyr)
library(reshape2)
library(kableExtra)
library(rstatix)

```



### Question 1 {-}

1. There are 30 candy covered chocolates in a bag M&M’s. There is a .1 probability that that the candy is red. If X is the number of red M&M’s in the bag.

i. Give the binomial probability mass function for X.

__ANSWER:__

\[\Pr(k)= \left( \begin{array}{c}
	30  \\
	k  \\
	\end{array} \right)(0.1)^k(0.9)^{30-k}, \ \ k=0,1,2,...30, \]
	where k is the number of red M & Ms.
	

ii. Find the probability of less than 2 red M&Ms in the bag.
 
__ANSWER:__


```{r}
n<-30
k<-1
MandM<-0:n # Number of Games
Pr_Red<-0.1 # Probability of Success
Pr_Reds<-c(dbinom(0:n,n,Pr_Red)) # Binomial Probability 
Q1ii<-c(dbinom(0:k,n,Pr_Red))
E_X<- n*Pr_Red # Expected Value
Var_X<-n*Pr_Red*(1-Pr_Red) # Variance

```

\[\Pr(0)= \left( \begin{array}{c}
	30  \\
0  \\
	\end{array} \right)(0.1)^0(0.9)^{30-0}= `r Q1ii[1]`\]
\[\Pr(1)= \left( \begin{array}{c}
	30  \\
1  \\
	\end{array} \right)(0.1)^1(0.9)^{30-1}= `r Q1ii[2]`\]

\[Pr(<2)=Pr(0)+Pr(1)=`r Q1ii[1]`+`r Q1ii[2]`=`r sum(Q1ii)`,  \]

As it is a Binomial Distribution we can state that the expected number of Red is
\[E[k]=`r n`\times `r Pr_Red` =`r E_X`,\]
the variance of the distribution is 
\[Var[k]=`r n`\times `r Pr_Red`\times(1-`r Pr_Red`) =`r Var_X`.\]

The plot below shows the Binomial Distribution M & Ms:

```{r}
df <- data.frame(MandM,Pr_Reds)
## Plot
binomial.p_dist<-ggplot(df, aes(x=MandM,y=Pr_Reds)) + geom_col(width = 0.1)+
  xlab("Number of Red M and Ms  out of 30")+
  ylab("Pr of Red")+ggtitle("Binomial Distribution")+
  scale_x_continuous(breaks=0:n)+ 
  geom_vline(xintercept = E_X, color = "Red", size=1)+
  geom_text(aes(x=E_X, label="E[k]", y=max(Pr_Reds)),vjust=2, 
            colour="red",  text=element_text(size=18))



 r<-c(-1,MandM) # Number of Games
CumPr_Reds<-c(0,cumsum(Pr_Reds))
df <- data.frame(r,CumPr_Reds)
  binomial.p_cum<-ggplot(df, aes(x=r,y=CumPr_Reds)) + geom_step()+
  xlab("r")+ylab("F(r)")+
  ggtitle("Cumulative Distribution")
  
ggarrange(  binomial.p_dist, binomial.p_cum, 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)

```



### Question 2 {-}

2. A baby wakes on average 0.25 times every hour.
 
i. If X is the number of times a baby wakes in an hour, give the poisson probability mass function for X.


__ANSWER:__

The distribution is described by the average, $\lambda=0.25$,
\[\Pr(k)=\frac{0.25^ke^{-0.25}}{k!}, \ \ k=0,1,2,... \]
where $k$ is the number of times the baby wakes every hour.

The expected value of the Poisson Distribution is
\[E[k]=0.25, \]
with the variance
\[Var[k]=0.25. \]



The plot below shows the Poisson Distribution for $\lambda=0.25$ average number of times a baby wakes in an hour:

```{r}
Wake<-0:20 # Number of Games
Baby_wakes<-0.25 # Lambda
Pr_wakes<-c(dpois(Wake,Baby_wakes)) # Poisson Probabilities
E_X<-Baby_wakes # Expected Outcome

df <- data.frame(Wake,Pr_wakes)

## Plot
Poisson.p_dist<-ggplot(df, aes(x=Wake,y=Pr_wakes)) + 
  geom_col(width = 0.1)+xlab("Number of Times a Baby wakes in an hour")+
  ylab("Pr of baby Wakes")+ggtitle("Poisson Distribution")+ 
  geom_vline(xintercept = E_X,  color = "blue", size=1)+
  geom_text(aes(x=E_X, label="E[k]", y=max(Pr_wakes)),vjust=2,
            colour="blue",text=element_text(size=18))


 r<-c(-1,Wake) # Number of Games
CumPr_wakes<-c(0,cumsum(Pr_wakes))
df <- data.frame(r,CumPr_wakes)
  Poisson.p_cum<-ggplot(df, aes(x=r,y=CumPr_wakes)) + geom_step()+
  xlab("r")+ylab("F(r)")+
  ggtitle("Cumulative Distribution")
  
ggarrange( Poisson.p_dist, Poisson.p_cum, 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)
```






ii. If X is the number of times a baby wakes in eight hour, give the poisson probability mass function for X.

__ANSWER:__

The distribution is described by the average, $\lambda=0.25\times 8 =2$,
\[\Pr(k)=\frac{2^ke^{-2}}{k!}, \ \ k=0,1,2,... \]
where $k$ is the number of times the baby wakes every 8 hours.

The expected value of the Poisson Distribution is
\[E[k]=2, \]
with the variance
\[Var[k]=2. \]


The plot below shows the Poisson Distribution for $\lambda=2$ average number of times a baby wakes in eight hours:

```{r}

Baby_wakes<-0.25*8 # Lambda
Pr_wakes<-c(dpois(Wake,Baby_wakes)) # Poisson Probabilities
E_X<-Baby_wakes # Expected Outcome

df <- data.frame(Wake,Pr_wakes)

## Plot
Poisson.p_dist<-ggplot(df, aes(x=Wake,y=Pr_wakes)) + 
  geom_col(width = 0.1)+xlab("Number of Times a Baby wakes in an hour")+
  ylab("Pr of baby Wakes")+ggtitle("Poisson Distribution")+ 
  geom_vline(xintercept = E_X,  color = "blue", size=1)+
  geom_text(aes(x=E_X, label="E[k]", y=max(Pr_wakes)),vjust=2,
            colour="blue",text=element_text(size=18))



 r<-c(0,Wake) # Number of Games
CumPr_wakes<-c(0,cumsum(Pr_wakes))
df <- data.frame(r,CumPr_wakes)
  Poisson.p_cum<-ggplot(df, aes(x=r,y=CumPr_wakes)) + geom_step()+
  xlab("r")+ylab("F(r)")+
  ggtitle("Cumulative Distribution")
  
ggarrange( Poisson.p_dist, Poisson.p_cum, 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)
```






iii. What is the probability that the baby does not wake during the 8 hours.

__ANSWER:__

```{r,echo=FALSE, warning=FALSE,dpi=300,echo=TRUE, warning=FALSE}

Baby_wakes<-0.25*8 # Lambda
Pr_wakes_zero<-dpois(0,Baby_wakes) #
```

\[\Pr(0)=\frac{2^0e^{-2}}{0!}=`r Pr_wakes_zero`.  \]

### Question 3 {-}

3. Give the features of a
 
i. Geometric Experiment. 

__ANSWER:__

* The experiment consists of a series of repeated Bernoulli trials

* There are two possible outcomes arbitrary called success and failure 

*  A success occurs with probability p and a failure occurs with probability q=1-p

*	The Random Variable is ordered as 1 if success and 0 if failure

*	The random variable is the number of trials performed to yield one success

* $E[X]=1/p$, $VAR[X]=q/p^2$


ii. Binomial Experiment. 

__ANSWER:__


* The experiment consists of n repeated Bernoulli trials

* The trials are independent

* The probability of success in each trial is constant

* E[x]=np, VAR[X]=npq



iii. Poisson Experiment.

__ANSWER:__

* The experiment consists of a number of events happening randomly over time or space

* The events occur independently of each other

* The rate of occurrence of the events is a well defined average per unit/space

* The Random Variable is the number of events occurring in a given interval

* E[X]=lambda, VAR[X]=lambda


iv. Negative Binomial Experiment.

__ANSWER:__

* The trials are independent

* The number of trials to be performed is not known at the start of the experiment 

* The probability of success in each trial is constant

* The random variable is the number of trials performed to yield $r$ successes
* $E[X]=r/p$, $VAR[X]=r\frac{q}{p^2}$

### Question 4 {-}

4. Every day a production line makes 100 computers of which 10% are defective. If X is the number of defective computers in a day.

i. Give the binomial probability mass function for X.

__ANSWER:__

\[\Pr(k)= \left( \begin{array}{c}
	100  \\
	k  \\
	\end{array} \right)(0.1)^k(0.9)^{100-k}, \ \ k=0,1,2,...100, \]
	where k is the number of defective computers.
	

ii. Find the probability that there is more than 2 computers defective in a day.


__ANSWER:__


```{r}
n<-100
k<-2
Computers<-0:n # Number of Games
Pr_Def<-0.1 # Probability of Success
Pr_Defs<-c(dbinom(0:n,n,Pr_Def)) # Binomial Probability 
Q4ii<-c(dbinom(0:k,n,Pr_Def))
E_X<- n*Pr_Def # Expected Value
Var_X<-n*Pr_Def*(1-Pr_Def) # Variance

```

\[\Pr(0)= \left( \begin{array}{c}
	100  \\
0  \\
	\end{array} \right)(0.1)^0(0.9)^{100-0}= `r Q4ii[1]`,\]
\[\Pr(1)= \left( \begin{array}{c}
	100  \\
1  \\
	\end{array} \right)(0.1)^1(0.9)^{100-1}= `r Q4ii[2]`,\]

\[Pr(2)= \left( \begin{array}{c}
	100  \\
2  \\
	\end{array} \right)(0.1)^2(0.9)^{100-2}= `r Q4ii[3]` .\]

\[ Pr(>2)=1-Pr(\leq2)=1-(Pr(0)+Pr(1)+Pr(2))=1-`r Q4ii[1]`+`r Q4ii[2]`+`r Q4ii[3] `  \]

\[ Pr(>2)= `r 1-sum(Q4ii)`  \]

iii. What is the E[X] and Var[X] of the distribution?

__ANSWER:__

As it is a Binomial Distribution we can state that the expected number  is
\[E[k]=`r E_X`,\]
the variance of the distribution is 
\[Var[k]=`r Var_X`.\]

The plot below shows the Binomial Distribution and the Cumulative for defective computers:

```{r}
df <- data.frame(Computers,Pr_Defs)
## Plot
binomial.p_dist<-ggplot(df, aes(x=Computers,y=Pr_Defs)) + geom_col(width = 0.1)+
  xlab("Number of Defective Computers")+
  ylab("Pr of Defective")+ggtitle("Binomial Distribution")+
  geom_vline(xintercept = E_X, color = "Red", size=1)+
  geom_text(aes(x=E_X, label="E[k]", y=max(Pr_Defs)),vjust=2, 
            colour="red",  text=element_text(size=18))


 r<-c(1,Computers) # Number of Games
CumPr_Defs<-c(0,cumsum(Pr_Defs))
df <- data.frame(r,CumPr_Defs)
  binomial.p_cum<-ggplot(df, aes(x=r,y=CumPr_Defs)) + geom_step()+
  xlab("r")+ylab("F(r)")+
  ggtitle("Cumulative Distribution")
  
ggarrange( binomial.p_dist, binomial.p_cum, 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)
 
```


### Question 5 {-}

5. A phone center receives 15 calls every 30 minutes.
 
i. If X is the number of phone calls in 30 minutes, give the Poisson probability mass function for X.

__ANSWER:__

The distribution is described by the average, $\lambda=15$,
\[\Pr(k)=\frac{15^ke^{-15}}{k!}, \ \ k=0,1,2,... \]
where $k$ is the number of phone calls per half hour.

The expected value of the Poisson Distribution is
\[E[k]=15, \]
with the variance
\[Var[k]=15. \]



The plot below shows the Poisson Distribution and the Cumulative Distribution for $\lambda=15$ average number of calls per half-hour:

```{r}
Calls<-0:50 # Number of Games
No_of_Calls<-15 # Lambda
Pr_Calls<-c(dpois(Calls,No_of_Calls)) # Poisson Probabilities
E_X<-No_of_Calls # Expected Outcome

df <- data.frame(Calls,Pr_Calls)

## Plot
Poisson.p_dist<-ggplot(df, aes(x=Calls,y=Pr_Calls)) + 
  geom_col(width = 0.1)+xlab("Number of Phone Calls in a half hour")+
  ylab("Pr of Phone Calls")+ggtitle("Poisson Distribution")+ 
  geom_vline(xintercept = E_X,  color = "blue", size=1)+
  geom_text(aes(x=E_X, label="E[k]", y=max(Pr_Calls)),vjust=2,
            colour="blue",text=element_text(size=18))

r<-c(-1,Calls) # Number of Games
CumPr_Calls<-c(0,cumsum(Pr_Calls))
df <- data.frame(r,CumPr_Calls)
  Poisson.p_cum<-ggplot(df, aes(x=r,y=CumPr_Calls)) + geom_step()+
  xlab("r")+ylab("F(r)")+
  ggtitle("Cumulative Distribution")


ggarrange( Poisson.p_dist, Poisson.p_cum, 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)
```



ii. What is the probability that there will be exactly 10 phone calls in the first 30 minutes and exactly 20 phone calls in the second 30 minutes.

__ANSWER:__


```{r}
No_of_Calls<-15 # Lambda
Pr_Calls_10<-dpois(10,No_of_Calls) #  10 Calls Poisson Probability
Pr_Calls_20<-dpois(20,No_of_Calls) #  20 Calls Poisson Probability
Pr_Calls_10_then_20 <- Pr_Calls_10*Pr_Calls_20
```

Probability of 10 calls in the first half-hour:
\[\Pr(10)=\frac{15^{10}e^{-15}}{10!}=`r Pr_Calls_10`. \]
Probability of 20 calls in the first half-hour:

\[\Pr(20)=\frac{15^{20}e^{-15}}{20!}=`r Pr_Calls_20`. \]

The combination of both:
\[\Pr(10)\times \Pr(20)=`r Pr_Calls_10_then_20`. \]

iii. What is the E[X] and Var[X] of the distribution.

__ANSWER:__

\[E[X]=\lambda= `r No_of_Calls` \]
\[Var[X]=\lambda= `r No_of_Calls` \]


A basketball player has a 0.3 chance of making a free throw. They keep shooing until they get a basket.

i. Give the appropriate distribution for this situation.
__ANSWER:__





This scenario describes repeated independent trials until the first success, which is a **Geometric Distribution**

The probability that the first success occurs on the \( k \)-th trial in a geometric distribution is:

\[
P(X = k) = (0.7)^{k - 1} \cdot 0.3
\]

where k=1,2,3,..., is the number of free throws until a success

ii. What is the probability they make their first shot on the third attempt?

__ANSWER:__


\[
P(X = 3) = (0.7)^2 \cdot 0.3 = 0.49 \cdot 0.3 = 0.147
\]

iii. What is the E[X] and Var[X] of the distribution.

__ANSWER:__

For a geometric distribution:

- **Expected Value**:
  \[
  E[X] = \frac{1}{p} = \frac{1}{0.3} = \boxed{3.\overline{3}}
  \]

- **Variance**:
  \[
  \text{Var}[X] = \frac{1 - p}{p^2} = \frac{0.7}{0.09} \approx \boxed{7.78}
  \]



### Question 7 {-}

A city installs a sensor at a traffic light that records the number of cars passing through every minute. On average, 3.3 cars pass through per minute.

i. Give the appropriate distribution for this situation.




__ANSWER:__

This scenario involves counting the number of events (cars passing) in a fixed interval (1 minute), with events occurring independently which is a **Poisson Distribution**

\[
P(X = k) = \frac{e^{-3.3}3.3^k}{k!}
\]
where k=0,1,2,... is the number of cars in a minute 


ii. What is the probability that fewer than 2 cars pass through in a minute?

__ANSWER:__

\[
P(X < 2) = P(X = 0) + P(X = 1)
\]



- \( P(X = 0) = \frac{e^{-3.3} \cdot 3.3^0}{0!} = e^{-3.3} = 0.0365 \)
- \( P(X = 1) = \frac{e^{-3.3} \cdot 3.3^1}{1!} = 3.3 \cdot e^{-3.3} = 0.1205 \)

\[
P(X < 2) = 0.0365 + 0.1205 = 0.157
\]

iii. What is the expected number of cars in a 10-minute interval?

__ANSWER:__


\[
E[X_{10}] = \lambda \cdot 10 = 3.3 \cdot 10 = 33.
\]



## Multiple-Choice Questions {-}

### MCQ Question 8 {-}

A multiple-choice quiz consists of ten questions each with five possible answers of which only one is correct.
What is the appropriate probability distribution.

i. Geometric Distribution; 

ii. Binomial Distribution; 

iii. Poisson Distribution;

iv. Negative Binomial Distribution;

v. Gaussian  Distribution.


__ANSWER:__

Binomial Distribution.

### MCQ Question 9 {-}

When a person fishing catches a fish, it is too small with a probability of 0.42 and it is returned to the water. On the other hand if it is bigger the person stops fishing.

i. Geometric Distribution; 

ii. Binomial Distribution; 

iii. Poisson Distribution;

iv. Negative Binomial Distribution;

v. Gaussian  Distribution.

__ANSWER:__

Geometric Distribution

### MCQ Question 10 {-}

The Poisson distribution is often used to model which of the following scenarios?

A) The number of events occurring within a fixed interval of time or space. 

B) The number of successes in a fixed number of independent trials.

C) The number of trials required to get the first success.

D) The number of failures before the first success in a series of Bernoulli trials.

__ANSWER:__

A 


### Question 11 {-}

Write your own question with a named discrete distribution.


